---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read functions for screeningmodels and screening plots
```{r}
source("sourcecode_screening2020.R")
```

Figure 8: Hierarchical generalised additive model for the region of Värmland
```{r}
Series<-read_tsv("Vdr 88-17 ANC300.txt")
Series%>%
  filter(SWEREFID=="6559832-380439" | SWEREFID=="6581519-383565" | SWEREFID=="6585489-420879" | SWEREFID=="6586310-399589" | SWEREFID=="6591038-411956"| SWEREFID=="6635190-525485" |
           SWEREFID=="6538630-449064" | SWEREFID=="6585343-440992"| SWEREFID=="6652237-421507")%>%
  mutate(Sulfate=`SO4* mekv/l`, date = paste(Yr, Mo, Day, sep="-") %>% 
           ymd())%>%
  mutate(datedecimal=decimal_date(date))->Series_V
```

Hierarchical model:
```{r}
Sulfate_modGS <- gam(Sulfate ~ s(datedecimal, k=5, m=2) + s(datedecimal, as.factor(SWEREFID) , k=5, bs="fs", m=2), data=Series_V, method="REML")
```

Make predictions to use in plot for overall trend:
```{r}
newdata<-data.frame(Sulfate=Series_V$Sulfate, SWEREFID="", datedecimal=Series_V$datedecimal)
newdata$overall<-predict(Sulfate_modGS, newdata=newdata, type="iterms", se.fit=TRUE)
newdata$overallfit<-newdata$overall$fit[,1]
newdata$overall.se<-newdata$overall$se.fit[,1]
```

Plot overall trend estimate with uncertainty (Figure 8, left)
```{r}
newdata%>%
  ggplot(aes(x=datedecimal, y=overallfit+Sulfate_modGS$coefficients[1]))+
  geom_point()+
  geom_ribbon(aes(ymin=overallfit+Sulfate_modGS$coefficients[1]-2*overall.se,
                  ymax=overallfit+Sulfate_modGS$coefficients[1]+2*overall.se), alpha=0.25, lwd=2)+
  ylab("Overall trend in sulfate")+
  xlab("Year")+
  ylim(0,0.15)+
  theme_bw()+
  theme(text = element_text(size=15))
```

Make prediction for spline deviation of specific sites

```{r}
newdata2<-data.frame(Sulfate=Series_V$Sulfate, SWEREFID=Series_V$SWEREFID, datedecimal=Series_V$datedecimal)
newdata2$overall<-predict(Sulfate_modGS, newdata=newdata2, type="terms")
```

Plot individual trends for specific sites (Figure 8, right):
```{r}
newdata2%>%
  ggplot(aes(x=datedecimal, y=overall[,2]+newdata$overall$fit[,1]+Sulfate_modGS$coefficients[1], col=as.factor(SWEREFID)))+
  geom_point()+
  ylab("Site-specific trends")+
  xlab("Year")+
  labs(col = "Site")+
  ylim(0,0.15)+
  theme_bw()+
  theme(legend.position="bottom", text = element_text(size=18))
```


Trend screening models for Sulfate, pH, TOC, absorbance, alkalinity, ANC, BC, Chloride, Nitrite and nitrate and total phosphorus

```{r}
Series_V%>%
  rename(TOC=`TOC (mg/l)`, Absorbance=`Abs._F (420nm/5cm)`, Alkalinity=`Alk/Acid (mekv/l)`, ANC=`ANC mekv/l`, BC=`BC* mekv/l`, Chloride=`Cl (mekv/l)`, `Nitrite and nitrate`=`NO2+NO3-N (?g/l)`, `Total phosphorus`=`Tot-P (?g/l)`) %>% 
  select(Sulfate,pH, TOC, Absorbance, Alkalinity, ANC, BC, Chloride, `Nitrite and nitrate`, `Total phosphorus`,  date , SWEREFID)%>%
  drop_na(Sulfate) %>% 
  pivot_longer(cols=c("Sulfate", "pH", "TOC", "Absorbance", "Alkalinity", "ANC", "BC", "Chloride", "Nitrite and nitrate", "Total phosphorus")) %>% 
  group_by(SWEREFID,name) %>% 
  mutate(n=length(date)) %>% 
  screeningmodeling(values=  value,
                    datevar = date, 
                    #seasonal = F, # removed, it is a bad idea
                    link = "identity", 
                    conf.type = "conf",
                    beep = TRUE, 
                    tdist = F,
                    autocor = TRUE,
                    SWEREFID, name) -> 
  screenmodel_V_out
```

Screening plots for the region of Värmland (Figure 9)

```{r}
screenmodel_V_out%>%
  plot_screeningtrends(sorting = name, y_id = name, wrappingvar = SWEREFID)

```

